Modeling thermal dust emission with two components: application to the {\it PLANCK} HFI maps

### 2014-07-24 ms.tex
abstract rewrite, this is the version sent around for feedback/comments

We apply the \cite{FDS99} two-component thermal dust emission model to the
\PLANCK~HFI maps. This parametrization of the far-infrared dust spectrum as the
sum of two modified blackbodies serves as an important alternative to the 
commonly adopted single modified blackbody (MBB) dust emission model. Analyzing
the joint \PLANCK/DIRBE dust spectrum, we show that two-component models 
provide a better fit to the 100-3000 GHz emission than do single-MBB models, 
though by a lesser margin than found by \cite{FDS99} based on FIRAS and DIRBE. 
We also derive full-sky $6.1'$ resolution maps of dust optical depth and 
temperature by fitting the two-component model to the \PLANCK~217, 353, 545, 
and 857 GHz maps along with  DIRBE/\IRAS~100$\mu$m data. Because our 
two-component model matches the dust spectrum near its peak, accounts for the 
flattening of the dust spectrum at millimeter wavelengths, and specifies dust 
temperature at 6.1$'$ FWHM, our model provides reliable, high-resolution 
thermal dust emission foreground predictions from 100 GHz to 3000 GHz. We find 
that, in diffuse sky regions, our two-component 100-217 GHz thermal dust 
emission predictions are on average accurate to within 2.2\%, which represents 
an improvement of 5-15\% relative to predictions based on the \cite{planckdust}
single-MBB model. We calibrate our two-component optical depth to reddening at 
visible wavelengths, and compare the resulting reddening estimates with those 
based on stellar spectra. We find imperfect zodiacal light subtraction in both 
DIRBE/\IRAS~and \PLANCK~to be the dominant large-scale systematic problem 
in our two-component temperature map and reddening estimates.

@@

Thus, our degree-resolution goodness-of-fit analysis has generally confirmed
the conclusions of $\S$\ref{sec:global}. We find the single-MBB $\beta$ value 
favored by the combination of \PLANCK~and DIRBE to be nearly identical here 
($\beta$=1.57) versus in $\S$\ref{sec:global} ($\beta$=1.59). As in 
$\S$\ref{sec:global}, we also find that the \PLANCK+FIRAS and \PLANCK+DIRBE 
best-fit two-component models from Table \ref{tab:global} outperform single-MBB
alternatives, though only by a relatively small margin. The agreement between 
our correlation slope analysis and the present goodness-of-fit analysis is 
especially encouraging for two main reasons: (1) in the present analysis, dust 
temperature has been allowed to vary on degree scales, whereas in 
$\S$\ref{sec:global} we assumed a single global dust temperature (2) in the
present analysis, our zero-level offsets factor into the dust temperature,
whereas in $\S$6 this was not the case, meaning the former and latter analyses 
agree in spite of their potential to be affected by rather different 
systematics.

@@

In both the correlation slope analysis of $\S$\ref{sec:global} and the 
goodness-of-fit analysis of $\S$\ref{sec:hier} we found that the 
best-fit \PLANCK+DIRBE two-component model (Table \ref{tab:global}, model 2) 
outperformed the best-fit single-MBB model, but by a lesser margin than found 
by FDS99 using FIRAS+DIRBE. Specifically, our best-fit \PLANCK+DIRBE 
two-component model yielded an improvement of $\Delta \chi^2_{\nu}$=3.41 
($\S$\ref{sec:global}) and $\Delta \chi^2_{\nu}$=0.4 ($\S$\ref{sec:hier}). This
represents a far less dramatic contrast than found by the FDS99 correlation 
slope analysis, $\Delta \chi^2_{\nu}$=29.2. Perhaps, though, the relative lack 
of discrimination amongst competing dust SED models when relying on 
\PLANCK+DIRBE is not surprising, given that our constraints include only nine 
broad frequency channels, whereas FDS99 employed $>$200 narrow bands.

@@

Nevertheless, we have established that the two-component emission model
remains viable in light of the \PLANCK~HFI data, and that the FIR/submm 
dust SED's preference for two MBB components rather than just one is not
simply an idiosyncracy of the FIRAS spectra. Furthermore, we showed in 
$\S$\ref{sec:lofreq} that our 100-217 GHz two-component emission predictions
are more accurate than those of \cite{planckdust} by margins of $\sim$5-15\%.
We therefore recommend that those interested in thermal dust foregrounds in 
the 100-3000 GHz frequency range use our data release to predict unpolarized 
dust emission, at the very least in order to help determine the level at which 
the choice of dust emission model may influence their conclusions.
%last sentence needs work

>>>>

We apply the \cite{FDS99} two-component thermal dust emission model to the
\PLANCK~HFI maps. This parametrization of the far-infrared dust spectrum as the
sum of two modified blackbodies serves as an important alternative to the 
commonly adopted single modified blackbody (MBB) dust emission model. Analyzing
the joint \PLANCK/DIRBE dust spectrum, we show that two-component models 
provide a better fit to the 100-3000 GHz emission than do single-MBB models, 
though by a lesser margin than found by \cite{FDS99} based on FIRAS and DIRBE. 
We also derive full-sky $6.1'$ resolution maps of dust optical depth and 
temperature by fitting the two-component model to \PLANCK~217, 353, 545, and 
857 GHz along with  DIRBE/\IRAS~100$\mu$m data. Because our two-component model
matches the dust spectrum near its peak, accounts for the flattening of the 
dust spectrum at millimeter wavelengths, and specifies dust temperature at 
6.1$'$ FWHM, our model provides reliable, high-resolution thermal dust emission
foreground predictions from 100 GHz to 3000 GHz. We find that, in diffuse sky 
regions, our two-component 100-217 GHz thermal dust emission predictions are on
average accurate to within 2.2\%, while the \cite{planckdust} single-MBB model 
systematically underpredicts emission by 18.8\% at 100 GHz, 12.6\% at 143 GHz 
and 7.9\% at 217 GHz. We calibrate our two-component optical depth to reddening
at visible wavelengths, and compare the resulting reddening estimates with 
those based on stellar spectra. We find imperfect zodiacal light subtraction in
both DIRBE/\IRAS~and \PLANCK~to be the dominant large-scale systematic problem 
in our two-component temperature map and reddening estimates.

@@

Thus, our degree-resolution goodness-of-fit analysis has generally confirmed
the conclusions of $\S$\ref{sec:global}. We find the single-MBB $\beta$ value 
favored by the combination of \PLANCK~and DIRBE to be nearly identical here 
($\beta$=1.57) versus in $\S$\ref{sec:global} ($\beta$=1.59). As in 
$\S$\ref{sec:global}, we also find that the \PLANCK+FIRAS and \PLANCK+DIRBE 
best-fit two-component models from Table \ref{tab:global} outperform single-MBB
alternatives, though only by a relatively small margin in $\chi^2_{\nu}$. 
Still, because our present analysis has $\sim$75,500 degrees-of-freedom, 
$\Delta \chi^2_{\nu}$=0.4 formally corresponds to an enormously significant
improvement in $\chi^2$. The agreement between our correlation slope analysis 
and the present goodness-of-fit analysis is especially encouraging for two main
reasons: (1) in the present analysis, dust temperature has been allowed to vary
on degree scales, whereas in $\S$\ref{sec:global} we assumed a single global 
dust temperature and (2) in the present analysis, our zero-level offsets factor
into the dust temperature, whereas in $\S$6 this was not the case, meaning the 
former and latter analyses agree in spite of their potential to be affected by 
rather different systematics.

@@

In both the correlation slope analysis of $\S$\ref{sec:global} and the 
goodness-of-fit analysis of $\S$\ref{sec:hier} we found that the 
best-fit \PLANCK+DIRBE two-component model (Table \ref{tab:global}, model 2) 
outperformed the best-fit single-MBB model, but by a lesser margin in 
$\chi^2_{\nu}$ than found by FDS99 using FIRAS+DIRBE. Specifically, our 
best-fit \PLANCK+DIRBE two-component model yielded an improvement of 
$\Delta \chi^2_{\nu}$=3.41 ($\S$\ref{sec:global}) and $\Delta \chi^2_{\nu}$=0.4
($\S$\ref{sec:hier}). This represents a far less dramatic contrast in 
$\chi^2_{\nu}$ than found by the FDS99 correlation slope analysis, 
$\Delta \chi^2_{\nu}$=29.2. Perhaps a relative lack of discrimination amongst 
competing dust SED models when relying on \PLANCK+DIRBE is to be expected, 
given that our constraints include only nine broad frequency channels, whereas 
FDS99 employed $>$200 narrow bands. Still, $\Delta \chi^2_{\nu}$=0.4 from 
$\S$\ref{sec:hier} is formally of enormous significance, given the 
$\sim$75,000 degrees of freedom in that analysis.

@@

Nevertheless, we have established that the two-component emission model
remains viable in light of the \PLANCK~HFI data, and that the FIR/submm 
dust SED's preference for two MBB components rather than just one is not
simply an idiosyncracy of the FIRAS spectra. Furthermore, we showed in 
$\S$\ref{sec:lofreq} that our 100-217 GHz two-component emission predictions
are on average accurate to within 2.2\%, whereas the \cite{planckdust} 
single-MBB model systematically understimates low-frequency dust emission by 
18.8\% at 100 GHz, 12.6\% at 143 GHz and 7.9\% at 217 GHz. We therefore 
recommend that those interested in thermal dust foregrounds in the 100-3000 GHz
frequency range use our data release to predict unpolarized dust emission, at 
the very least in order to help determine the level at which the choice of dust
emission model may influence their conclusions.
%last sentence needs work
